
\documentclass[a4paper,letterpaper,12pt,oneside,draft]{article}

\usepackage{geometry}
\geometry{margin=2cm,hoffset=0in, %
    headheight=0.5\baselineskip}
\usepackage{times}
\pagestyle{plain}
\usepackage{setspace}
\onehalfspacing
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{listings}

\input{../setup.tex}

\newcommand{\eec}{\;,}
\newcommand{\eep}{\;.}
\newcommand{\zth}{0th }
\newcommand{\fst}{1st }
\newcommand{\snd}{2nd }
\newcommand{\OpL}{\mathscr{L}}
\newcommand{\OpT}{\mathscr{T}}
\newcommand{\MG}{MG }

\newcommand{\allspace}{\ensuremath{\mathbb{R}^3}}
\newcommand{\norm}[1]{\left| #1 \right|}
\newcommand{\bracket}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\bracketv}[1]{\bracket{#1}_V}
\newcommand{\bracketvo}[1]{\ensuremath{\bracket{#1}_{V,\Omega}}}
\newcommand{\bracketR}[1]{\ensuremath{\bracket{#1}_{\allspace}}}
\newcommand{\bracketex}[2][V]{\ensuremath{\bracket{#2}_{#1}}}
\newcommand{\bracketsa}[1]{\bracketex[\allspace,\Omega]{#1}}
\newcommand{\rdotJ}[1][\allspace]{\bracketex[#1]{\vec{r}\cdot\vec{J}}}
\newcommand{\rsqp}{\bracketR{r^2\phi}}
\newcommand{\intg}[2][g]{\ensuremath{\int_{E_{#1}}^{E_{#1-1}} #2 dE}}
\newcommand{\intcg}[2][g]{\ensuremath{\int_{E_{#1}}^{\infty} #2 dE}}
\newcommand{\vr}{\ensuremath{\vec{r}}}
\newcommand{\dvr}{\left(\vr-\vr_0\right)}
\newcommand{\psif}[1][]{\psi(\vr,\Omega#1,E#1)}
\newcommand{\psifz}[1][]{\psi(\vr,\Omega#1,E#1;\vr_0)}
\newcommand{\dvrdotJ}{\bracketR{\dvr\cdot\vec{J}}}
\newcommand{\dvrdotJg}{\bracketR{\dvr\cdot\vec{J_g}}}
\newcommand{\dvrsqp}{\bracketR{\dvr^2\phi}}
\newcommand{\dvrsqpg}{\bracketR{\dvr^2\phi_g}}
\newcommand{\regint}[1]{\ensuremath{\int_{V_0} #1 dV_0}}

\counterwithin{equation}{section}

\title{Possible generalizations to migration based diffusion coefficient definitions}
\author{Eshed Magali \\ Edward W. Larsen}
\date{\today}


\widowpenalty10000
\clubpenalty10000

\begin{document}
\maketitle
\section{Definitions}
In single assembly problems, the diffusion coefficients are chosen such that the mean migration distance of migration is preserved. 
Specifically, the diffusion coefficient is defined according to the following formula:
\begin{align}
    \nonumber
    D_g &\equiv \frac{\intg{\int_{\text{unitcell}} F(\vr_0)\int_{\allspace}\dvr^2 \Bigg[ \Sigma_t(\vr,E)\phi(\vr,E;\vr_0) - \int_0^\infty \Sigma_{s0}(\vr,E'\to E) \phi(\vr,E';\vr_0) dE'\Bigg]dVdV_0}} {6\intg{\int_{\text{unitcell}}F(\vr_0)\int_{\allspace} \phi(\vr,E;\vr_0)dVdV_0}} \\ \nonumber
    &= \frac{\intg{\int_{\text{unitcell}} F(\vr_0)\int_{\allspace}\OpL_0 \dvr^2 \phi(\vr,E;\vr_0) dVdV_0}} {6\intg{\int_{\allspace} \phi(\vr,E)dV}} \\
    \label{eq:SingleAssemblyDiffusion:def}
    &= \frac{\intg{\int_{\text{unitcell}} F(\vr_0)\int_{\allspace}\dvr\cdot\vec{J}(\vr,E;\vr_0) dVdV_0}} {3\intg{\int_{\allspace} \phi(\vr,E)dV}}\eec
\end{align}
where $F(\vr)$ is the fission source of the eigenvalue transport problem at $\vr$.

The equivalence of \cref{eq:SingleAssemblyDiffusion:def} to CMM as published by Liu et al.~\cite{Liu2018} appears in my Master's thesis~\cite{EshedThesis}. 
The numerator of \cref{eq:SingleAssemblyDiffusion:def} is the MC tally in CMM as of 2018~\cite{Liu2018}. 
The denominator is the group flux in the transport solution of the eigenvalue problem of the single assembly lattice.

This definition causes a preservation of the first moment of migration in the diffusion problem.
The diffusion problem for a single assembly lattice is homogeneous. 
In this case, from Fick's law one can derive for any point source in this medium:
\begin{equation}
    \vec{J}_g(\vr;\vr_0) = -D_g\nabla\phi_g(\vr;\vr_0)\eec
\end{equation}
and this equation can be multiplied by the strength of the diffusion fission source at $\vr_0$ and by $\dvr$ and integrated over all space for the $\vr$ variable and over the volume of the unit cell for $\vr_0$. 
$F(\vr)$ is constant in space because the solution of the homogeneous diffusion problem is spatially flat.
This results in the following form:
\begin{align}
    \nonumber
    \int_{\text{unitcell}} F(\vr_0)\int_{\allspace} \dvr\cdot\vec{J}_g(\vr;\vr_0) dVdV_0 &= -D_g \int_{\text{unitcell}} F(\vr_0) \int_{\allspace} \dvr\cdot\nabla \phi_g(\vr;\vr_0)dVdV_0 \\\nonumber
    &= 3D_g \int_{\text{unitcell}} F(\vr_0) \int_{\allspace}\phi_g(\vr;\vr_0)dVdV_0 \\
    \label{eq:SingleAssemblyDiffusion:Ficks}
    &= 3D_g \int_{\allspace}\phi_g(\vr)dV\eep
\end{align}
This means that if the integral flux spectrum is preserved between the transport and diffusion problem, which is guaranteed by the choice of the cross sections, then so will the first moment of migration.
This is a result of the combination of \cref{eq:SingleAssemblyDiffusion:def,eq:SingleAssemblyDiffusion:Ficks}.

The philosophy of this is that we try to preserve how far away neutrons stream from their point of origin, so that neutrons born in the diffusion problem travel on average similar distances from their origin.

When the diffusion problem is no longer infinite and homogeneous, there could be many ways to generalize this quantity.
First, there is the question of what the diffusion problem's fission source should look like in \cref{eq:SingleAssemblyDiffusion:Ficks}.
The only effect this has is on the exact interpretation of that function in that equation, and on the data needed to assign values to it.
Second, there is the question of whether we should preserve what neutrons born within a given region do, or if we should preserve the migration of neutrons that fly through a given cell, regardless of if they were born in it.

Lastly, we can attempt to define the neutron's behavior not with a homogenized colorset where each homogenized assembly is next to different homogenized assemblies but rather by driving an infinite homogeneous problem with the transport information.
This option might seem unphysical, because the homogenized problem is not simply a homogenization of the transport problem. 
However, it most accurately relates to the idea that the diffusion coefficient is a property of the material itself and not determined by other materials around it.

\subsection{Fission source distribution definition}
The main three options that we see are that one can use a homogeneous fission source, the transport problem's fission source or the resulting diffusion problem's fission source. 
Using a homogeneous fission source assumes that since the region is homogenized, the source in it should be treated as spatially flat too.
This choice is mostly compatible with CMM's results, but the last line in \cref{eq:SingleAssemblyDiffusion:Ficks} will no longer be true, as the fission source is no longer related to the actual fission source in the diffusion problem.
This will imply in most cases that Fick's law no longer relates the scalar flux to the first migration moment.
This is a major flaw, as one must then cause the preservation of a Green's function preservation rather than the scalar flux.
Cases where the relation only involves the scalar flux would still be considered later.

Defining the fission source within a region using the heterogeneous transport fission source asks the question "were we to simply change the physics of how neutrons move without changing their point of birth, what would that effect be".
However, this has two significant flaws.
First, the material in the diffusion problem is homogenized in space and in energy.
The physics has already changed by that, so the driving philosophy is already invalid, except for already-homogeneous problems.
Second, using the transport problem's fission source is not compatible with what CMM currently does.
CMM currently assigns a flat fission source (which is also the diffusion problem's eigenvalue fission source), and this change would require a change for single assembly problems as well.
Therefore we will not consider these options further.

Lastly, the fission source can be defined as the fission source determined by solving the multigroup homogenized diffusion problem.
This is an implicit mathematical definition.
The diffusion coefficient is theoretically defined after one already knows the distribution of the fission source in the diffusion problem, although this fission source distribution depends on the diffusion coefficient values.
This can, however, be treated in two separate ways.

First, one can use a definition where one does not actually need to know the exact value of this distribution.
This is done using Green's theory, through the transition used in the last line of \cref{eq:SingleAssemblyDiffusion:Ficks}.
Second, one can use a non-linear iteration to look for diffusion coefficients that cause the required preservation.
One acquires a non-linear equation that defines the first migration moment in terms of the diffusion coefficient and the scalar flux.
Then the diffusion coefficient is iterated until preservation of the transport first migration moment is achieved.

\subsubsection*{Naming}
When naming a diffusion coefficient definition, we call it either a homogeneous fission distributed method if the fission distribution is constant in space or a diffusion fission distributed method if the diffusion problem's eigenvalue fission source is used.

\subsection{Tallying by source region or by flight region}
In color set problems, neutrons can fly through regions in the transport problems that do not resemble their origin.
This raises the question, do we want to preserve for each region how far away neutrons that are born in that region migrate from their point of origin on average?
Or rather, do we want to preserve the migration neutrons do through a given region?

Mathematically, this has an effect on the region of integration.
In a single assembly problem we integrate the possible flight paths through all of space, and the fission source location over the entire unit cell, as seen in the LHS of \cref{eq:SingleAssemblyDiffusion:Ficks}.
When we consider a color set problem, we must obtain a different diffusion coefficient for each region. We must therefore limit our integrals so that we can possibly get a different diffusion coefficient for each region.
Therefore we can either limit the flight path integral, the source range integral, or both.
Limiting both makes little sense, in our opinion, as it means that we only count the flight paths neutrons have through a specific cell if they were also born in a cell of that type.
It will have the weaknesses of both methods and less strengths.

Limiting the source range would require tallying according to the following formula:
\begin{align}
    \nonumber
    T_{g,i} &= \intg{\int_{V_i}F(\vr_0)\int_{\allspace}\dvr^2\OpL_0\phi(\vr,E;\vr_0)dVdV_0} \\
    \label{eq:Tally:limitsource}
    &= \intg{\int_{V_i}F(\vr_0)\int_{\allspace}\dvr\cdot\vec{J}(\vr,E;\vr_0)dVdV_0} \eec
\end{align}
where $V_i$ is a single region of type $i$ in the infinite lattice of the transport problem.
This means that whenever a neutron path or interaction happens in any region, one counts it towards the tally of the neutron's ORIGIN's region.

Limiting the flight path would require tallying differently, according to the following formula:
\begin{align}
    \nonumber
    T_{g,i} &= \intg{\int_{\text{unitcell}}F(\vr_0)\int_{V_i^\infty} \dvr^2\OpL_0\phi(\vr,E;\vr_0)dVdV_0} \\
    \label{eq:Tally:limitflight}
    &= \intg{\int_{\text{unitcell}}F(\vr_0)\int_{V_i^\infty} \dvr\cdot\vec{J}(\vr,E;\vr_0)dVdV_0}\eec
\end{align}
where $V_i^\infty$ is the entire infinite volume of the full 3D infinite lattice that contains regions of type $i$.
This means that neutrons contribute to the tally of regions of type $i$ whenever they interact or move through a region of type $i$, regardless of their region of origin.

Using the second method, where tallies are assigned to the location of traversal, allows us to homogenize regions that have no internal fission source.
However, using this method will lead to a more complicated implementation for the definition of diffusion coefficients that includes surface terms.
These surface terms can result in a non-linear relation to define the diffusion coefficient, which would require an iterative search.
It can be circumvented by defining the diffusion coefficients using not only the diffusion flux but also including some surface tallies.
Even so, these surface tallies can be difficult to accurately obtain using Monte Carlo methods.

\subsubsection*{Naming}
We call a method either source region based if we group tallies according to the neutron's region of origin or flight region based if we group tallies according to the neutron's point of interaction (or flight path).

\subsection{Different homogenized problems for each region}
The most appealing philosophy, in our opinion, is when the homogenized diffusion problem being considered is the homogenized color set problem.
In this case the same homogenized problem is used to determine the multigroup cross sections and the diffusion coefficient, and the only difference is through which physical quantities they are chosen to preserve.
In a homogenized color set problem the spatial description is essentially a lattice where each cell is given a homogeneous color with homogeneous parameters, with no internal details.
However, this means that in the diffusion problem, and thus in the resulting definitions, the diffusion coefficients can depend not only on the flux spectrum within the homogenized cell but on the spectrum outside it as well.
This can happen with the introduction of surface terms, or with the tally of neutron flight paths through other regions.
One can potentially generalize CMM in a way that circumvents this altogether.

If for every cell we assume the homogenized diffusion problem that defines its diffusion coefficient is an infinite homogeneous problem, we will get a relatively simple derivation.
It doesn't make sense to include tallies from regions other than the neutron's internal source in this case, because they have no equivalent in the homogenized diffusion problem.
Therefore this is only truly compatible with tallying the neutrons by source region, and thus this is limited to cells with internal fission sources.

\subsubsection*{Naming}
When a definition uses infinite homogeneous diffusion problems we call it a separated infinite homogeneous definition.

\section{Picking appealing options}
The preferred fission source distribution in our opinion is the diffusion fission source, and we find the homogenized color set problem more appealing than a set of infinite homogeneous problems for each region. 
We are unsure about which tally scheme is preferable, and have different opinions on the matter.
In this section we specify the details for several methods that we find relatively appealing.
We start with the two we support most, and work our way down.

\subsection{A diffusion fission distributed, flight region based method}
In this case we tally $T_{g,i}$ according to \cref{eq:Tally:limitflight}, and use the diffusion fission distribution. \cref{eq:SingleAssemblyDiffusion:Ficks} works slightly different now:
\begin{align}
    \nonumber
    \int_{\text{unitcell}} F(\vr_0)\int_{V_i^\infty} \dvr\cdot\vec{J}_g(\vr;\vr_0) dVdV_0 &= -D_{g,i} \int_{\text{unitcell}} F(\vr_0) \int_{V_i^\infty} \dvr\cdot\nabla \phi_g(\vr;\vr_0)dVdV_0 \\\nonumber
    &= 3D{g,i} \int_{V_i^\infty}\phi_g(\vr)dV \\ 
    \label{eq:DiffFissFlight}
    &\quad -D{g,i}\int_{\text{unitcell}} F(\vr_0) \int_{\partial V_i^\infty}\phi_g(\vr;\vr_0)\dvr\cdot dAdV_0 \eep
\end{align}
The left hand side is nothing but the tally in \cref{eq:Tally:limitflight}, and can be considered to be a known quantity we want to obtain.
The right hand side includes both the familiar term with the scalar flux we make sure is preserved and a surface quantity. 
If we tally this quantity on each surface between assemblies with its own tally, we can plug in all these tallied quantities and solve easily for the diffusion coefficient.
Notice, however, that this tally can be difficult to calculate in Monte Carlo. 
This requires knowledge of the scalar flux on a boundary.

Another option exists. 
We can for any given diffusion coefficient calculate the diffusion value of this surface term, be it by directly solving the Green's function problem for the homogenized diffusion problem or by other means.
Once this quantity is obtained, it can be compared to the tallied left hand side, and we can iteratively search for a diffusion coefficient value that matches the two sides of \cref{eq:DiffFissFlight}.
This is what we attempted for the boundary source problem, but had convergence issues.

Notice that we have no a priori reason to know which method is better.
It has to be tested to see which definition is actually the correct one, if any.

\subsection{A diffusion fission distributed, source region based method}
In this case we tally $T_{g,i}$ according to \cref{eq:Tally:limitsource}, and use the diffusion fission distribution. \cref{eq:SingleAssemblyDiffusion:Ficks} works slightly different now:
\begin{align}
    \nonumber
    \int_{V_i} F(\vr_0)\int_{\allspace} \dvr\cdot\vec{J}_g(\vr;\vr_0) dVdV_0 &= -D_g \int_{V_i} F(\vr_0) \int_{\allspace} \dvr\cdot\nabla \phi_g(\vr;\vr_0)dVdV_0 \\
    \label{eq:DiffFissSource}
    &= 3D_g \int_{V_i}F(\vr_0)\int_{\allspace}\phi_g(\vr;\vr_0)dVdV_0 
\end{align}
Notice that the right hand side requires that we know the total fraction of the scalar flux that originates from within region $i$, but requires no surface terms.
The problem is, of course, that we need this fraction for the diffusion problem. 
We can assume that the total flux in each region is preserved from reaction rate preservation. 
However, the fraction of the flux that comes from each specific region is not preserved, and must either be calculated or tallied.
The same problems we had for the surface problems exist if we choose to have this calculated, because this depends on the diffusion coefficient and causes an implicit definition for the diffusion coefficient.
Unlike the surface terms from the previous definition, these terms are fairly easy to obtain in Monte Carlo.

Again, there is no reason to prefer one definition over the other, or those from the previous method.
Selection should depend on which is good enough in cases of interest and is implementable in practice.

\subsection{A homogenized fission distributed, source region based, separated infinite homogeneous method}
In this case we tally $T_{g,i}$ according to \cref{eq:Tally:limitsource}, and use a homogenized fission distribution.
This causes the Fick's law equation to be almost identical to that in \cref{eq:SingleAssemblyDiffusion:Ficks}. 
The only difference is in the regions of integration:
\begin{align}
    \nonumber
    \int_{V_i} F(\vr_0)\int_{\allspace} \dvr\cdot\vec{J}_g(\vr;\vr_0) dVdV_0 &= -D_g \int_{V_i} F(\vr_0) \int_{\allspace} \dvr\cdot\nabla \phi_g(\vr;\vr_0)dVdV_0 \\\nonumber
    &= 3D_g \int_{V_i} F(\vr_0) \int_{\allspace}\phi_g(\vr;\vr_0)dVdV_0 \\
    \label{eq:HomFissSourceInf}
    &= 3D_g \int_{\allspace}\phi_g(\vr)dV\eec
\end{align}
but because in the homogenized problem the homogenized fission distribution is appropriate, the equation is simplified just as it did before.
This means that the diffusion coefficient takes the same value it did in CMM, up to the necessary change in tallies such that they attribute interactions and flight paths to the origin's region.
We find this method less appealing, but we have no proof if it is flawed, and it should be fairly easy to test.

\section{Tabling the different options}

Each of these definitions has strengths and weaknesses. 
Some of them are compatible with single-assembly CMM, some are not.
These methods are compared in \cref{tbl:options}.

\begin{table}[h!]
    \caption{A comparison of different options for diffusion coefficient methodologies.}\label{tbl:options}
    \begin{tabular}{|c|c|c|c|}
        \hline
        Definition number & Strengths & Weaknesses & Mathematical properties \\
        \hline
        \#1 & CMM Compatible & Fission required & $F(\vr_0)$ is constant in diffusion \\
        & & Difficult implementation & Tally tracked by region of birth \\
        & & Unclear interpretation & See \cref{eq:def:1} \\
        \hline
        \#2 & & Source spectrum undefined & $F(\vr_0)$ is constant in diffusion \\
        & & Difficult implementation & Tally tracked by region of flight \\
        & & Unclear interpretation & See \cref{eq:def:2} \\
        \hline
        \#3 & CMM Compatible & Fission required & $F(\vr_0)$ is constant in diffusion \\
        & Easy implementation & & Tally tracked by region of birth \\
        & & & See \cref{eq:def:3} \\
        \hline
        \#4 & CMM Compatible & Fission required & $F(\vr_0)$ is constant in diffusion \\
        & Easy implementation & Source spectrum undefined & Tally tracked by region of flight \\
        & &  & See \cref{eq:def:4} \\
        \hline
        \#5 & & Fission required & $F(\vr_0)$ is taken directly from transport \\
        & & Difficult implementation & Tally tracked by region of birth \\
        & & Unclear interpretation & See \cref{eq:def:1} \\
        & & Not backward compatible & \\
        \hline
        \#6 & & Source spectrum undefined & $F(\vr_0)$ is constant in diffusion \\
        & & Difficult implementation & Tally tracked by region of flight \\
        & & Unclear interpretation & See \cref{eq:def:2} \\
        \hline
    \end{tabular}
\end{table}

\end{document}

\begin{comment}
\begin{enumerate}
\item If we were to drop a neutron randomly inside the homogenized region, we would want it to travel, on average, the same distance that neutrons born in the region traveled in the transport problem.
\item If we were to drop a neutron randomly inside the homogenized region, we would want it to travel, on average, the same distance that neutrons traveled in this region in the transport problem.
\item If we were to drop a neutron randomly in an infinite homogeneous space made up of the homogenized region, we would want it to travel, on average, the same distance that neutrons born in the region traveled in the transport problem.
\item If we were to drop a neutron randomly in an infinite homogeneous space made up of the homogenized region, we would want it to travel, on average, the same distance that neutrons traveled in this region in the transport problem.
\item If we were to drop a neutron according to the transport fission source inside the homogenized region, we would want it to travel, on average, the same distance that neutrons born in the region traveled in the transport problem.
\item If we were to drop a neutron according to the transport fission source inside the homogenized region, we would want it to travel, on average, the same distance that neutrons born in the region traveled in the transport problem.
\item Neutrons born in the diffusion problem according to the diffusion fission source should travel on average the same distance that neutrons traveled in this region in the transport problem.
\item The average distance neutrons migrate away from their origin (the first migration moment) covered by neutrons in the cell in the diffusion problem should be the same as in the transport problem, given only that the integral scalar flux energy distribution is preserved.
\item The average distance neutrons migrate away from their origin covered by neutrons in the cell in the diffusion problem should be the same as in the transport problem, given that the scalar flux distribution is preserved and that surface terms are preserved on each surface.
\item Neutrons who enter the homogenized region through birth or streaming are kept as a driving source of migration (each carries its original point of birth). This source is treated as an input source on an infinite homogeneous diffusion problem, and the diffusion coefficient is chosen so that on average the migration moment in the diffusion problem is the same as in the transport problem.
\end{enumerate}
\end{comment}